Portable Multi- and Many-Core Performance for Finite Difference Codes; Application to the Free-Surface Component of NEMO.

### 2017-09-14 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite2d-paper: complete first draught of new similar-approaches section

@@

Although {PS}y{KA}l is a design for architecting code, rather than an
implementation of a tool, there is an implicit assumption that the
majority of the kernels in an application will be provided by the
application developer. In the GridTools~\citep{grid_tools} and
Firedrake~\citep{firedrake,fenics} approaches, the mathematical
operations (finite-difference stencils and finite element operations,
respectively) are specified in a high-level language by the
user. Kernel code is then generated automatically. It is possible to
optimise this generated code~\citep{coffee} and, in the case of
GridTools, support both CPU and GPU architectures.

@@

Lower-level abstractions, e.g. OCCA.

H. C. Edwards, C. R. Trott, D. Sunderland, "Kokkos: Enabling manycore performance portability through polymorphic memory access patterns", J. Parallel Distrib. Comput., vol. 74, no. 12, pp. 3202-3216, 2014.

>>>>

This paper is concerned with the implications of {PS}y{KA}l as a
design for architecting code. The implementation of an associated tool
for generating the middle, {PS}y, layer will be the subject of a future
paper. However, comparison with other approaches necessarily involves
tools rather than simply architectures.

@@

In the {PS}y{KA}l approach there is an implicit assumption that the
majority of the kernels in an application will be provided by the
application developer. In the GridTools~\citep{grid_tools} and
Firedrake~\citep{firedrake,fenics} approaches, the mathematical
operations (finite-difference stencils and finite element operations,
respectively) are specified in a high-level language by the
user. Kernel code is then generated automatically. It is possible to
optimise this generated code~\citep{coffee} and, in the case of
GridTools, support both CPU and GPU architectures. This ability to
generate kernels specific to a given computer architecture is a
powerful feature. However, with this power comes the responsibility of
providing domain scientists with the necessary functionality to
describe all conceivable kernel operations as well as a programming
interface with which they are comfortable.

The {PS}y{KA}l, GridTools and Firedrake approaches are all based on
the concept of (various flavours of) a Domain-Specific Language (DSL)
for finite-difference and finite-element applications. This is
distinct from other, lower-level abstractions such as
Kokkos~\citep{kokkos} and OCCA~\citep{occa} where the aim is to
provide a language that permits a user to implement a kernel just once
and have it compile to performant code on a range of multi- and
many-core devices. One could imagine using such abstractions to
implement kernels for the {PS}y{KA}l, GridTools and Firedrake
approaches rather than using e.g. OpenMP or CUDA directly.

@@

\begin{table}
\caption{An overview of the functionality of similar approaches. Static compilation here means that all code is compiled before program execution is begun.}
\label{TABLE_dsl_compare}
\centering
\begin{tabular}{l|c|c|c|c|c|c|c}
\hline
Approach   & DSL & MPI & Threading & Data layout & Kernels & Language & Compilation  \\
\hline
{PS}y{KA}l & Yes & Yes & Yes & Fixed  & User-supplied & Fortran & Static \\
GridTools  & Yes & No  & Yes & Flexible & Generated & C++  & Static \\
Firedrake  & Yes & Yes & No  & Fixed  & Generated & C with Python interface & Run-time \\
Kokkos     & No  & No  & Yes & Flexible & User-supplied & C++ & Static \\
OCCA       & No  & No  & Yes & Fixed  & User-supplied & C (Python \& Fortran interfaces) & Run-time \\
\hline
\end{tabular}
\end{table}
